Bird beak's elaboration and innovation

### 2021-11-16 Analysis/03.1-results.Rmd
Update 03.1-results.Rmd

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE, results = 'hide', message = FALSE, warning = FALSE}
## Run the analyses between groups
super_results_between <- dispRity.covar.projections(superorder_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses between groups
order_results_between <- dispRity.covar.projections(order_shapespace,
                                             type = "groups",
                                             base = "phylogeny")
## Run the analyses of groups on the phylogeny
super_results_phylo <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            verbose = TRUE)
## Save all
shapespace_allbirds_lvl_superorder_order_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(shapespace_allbirds_lvl_superorder_order_results_list, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
```

@@

```{r "starting with boxplots", echo = FALSE, fig.height = 8, fig.width = 8}
## Plot the results
wrap.plot <- function(results, col, main, names, centre.position = TRUE, absolute = FALSE) {

    ## Adding one plot
    one.plot <- function(one_result, xlab, col, zero = TRUE, main, centre, absolute, ...) {
        values <- lapply(get.disparity(one_result, concatenate = FALSE), c)
        if(centre) {
            values <- lapply(values, function(X) {X-0.5})
        }
        if(absolute) {
            values <- lapply(values, function(X) {abs(X-0.5)})
        }
        boxplot(values, horizontal = TRUE, yaxt = "n", col = col, xlab = xlab, main = main, ...)
        if(zero) {
            abline(v = 0, col = "black", lty = 3, lwd = 1)
        }
    }

@@

    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    op <- par(mfrow = c(1, n+1))
    ## Plot the first panel
    par(bty = "n", mar = c(5, 14, 4, 0) + 0.1)
    empty <- matrix(0, ncol = length(results[[1]]$disparity))
    colnames(empty) <- names(results[[1]]$disparity)
    boxplot(empty, horizontal = TRUE, las = 2, col = "white", ylim = c(0,0), xaxt = "n", border = "white")
    for(i in 1:n) {
        ## Is position?
        is_position <- names(results)[[i]] == "position"
        par(mar = c(5, 0, 4, 0) + 0.1, bty = "o")
        one.plot(results[[i]], col = col, centre = is_position && centre.position, main = ifelse(i == centre, main, ""), xlab = names_list[i], zero = is_position, absolute = absolute)
    }
    par(op)
}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

@@

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

```{r, fig.height = 16, fig.width = 8, echo = FALSE}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot the results
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on phylogeny] superorders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on phylogeny] superorders species\nonto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

## Plot the results
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on phylogeny] orders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation", "angle"))
```

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r "hehehe", echo = FALSE, fig.height = 16, fig.width = 8, eval = FALSE}
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on phylogeny] orders\nspecies onto the phylogenetic VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
````
</details>

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE}

## Plot the results
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on group] superorders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r "beforebefore last", echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on group] superorders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

@@

```{r "before last", fig.height = 16, fig.width = 8, echo = FALSE}

## Plot the results
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on group] oreders species\nonto their group's VCV",
          names = c("elaboration", "innovation", "angle"))
```

<details>
  <summary>[Click to expand] The same but using absolute values for the projections (elaboration)
</summary>

```{r "done with boxplots", echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on group] oreders species\nonto their group's VCV",
          names = c("absolute elaboration", "innovation", "angle"),
          centre.position = FALSE, absolute = TRUE)
```
</details>

>>>>

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE, results = 'hide', message = FALSE, warning = FALSE}
## Run the analyses between groups
super_results_between <- dispRity.covar.projections(superorder_shapespace,
                                             type = "groups",
                                             base = "phylogeny",
                                             output = c("position", "distance"))
## Run the analyses between groups
order_results_between <- dispRity.covar.projections(order_shapespace,
                                             type = "groups",
                                             base = "phylogeny",
                                             output = c("position", "distance"))
## Run the analyses of groups on the phylogeny
super_results_phylo <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE,
                                            output = c("position", "distance"))
## Run the analyses of groups on the phylogeny
order_results_phylo <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            base = "phylogeny",
                                            inc.base = TRUE,
                                            verbose = TRUE,
                                            output = c("position", "distance"))
## Run the analyses groups on themselves
super_results_within <- dispRity.covar.projections(superorder_shapespace,
                                            type = "elements",
                                            verbose = TRUE,
                                            output = c("position", "distance"))
## Run the analyses groups on themselves
order_results_within <- dispRity.covar.projections(order_shapespace,
                                            type = "elements",
                                            verbose = TRUE,
                                            output = c("position", "distance"))
## Save all
shapespace_allbirds_lvl_superorder_order_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(shapespace_allbirds_lvl_superorder_order_results_list, file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order_results_list.rda")
```

@@

```{r "starting with boxplots", echo = FALSE, fig.height = 8, fig.width = 8}
## Plot the results
wrap.plot <- function(results, col, main, names, centre.position = TRUE, absolute = TRUE, order = TRUE, xlim = NULL) {

    ## Ranking the results
    if(order) {
        ## Get the ranking
        orders <- do.call(cbind, lapply(results, get.disparity, concatenate = FALSE))
        ## Get the mean values
        orders <- apply(orders, c(1,2), function(x) mean(abs(unlist(x))))
        ## Sum them and rank them
        rank <- order(apply(orders, 1, sum), decreasing = FALSE)
    } else {
        rank <- 1:length(results[[1]]$disparity)
    }

    ## Adding one plot
    one.plot <- function(one_result, xlab, col, zero = TRUE, main, centre, absolute, rank = NULL, xlim = NULL, ...) {
        values <- lapply(get.disparity(one_result, concatenate = FALSE), c)
        if(centre) {
            values <- lapply(values, function(X) {X-0.5})
        }
        if(absolute) {
            values <- lapply(values, function(X) {abs(X-0.5)})
        }
        if(!is.null(rank)) {
            values <- values[rank]
        }
        if(!is.null(xlim)) {
            xlim <- xlim
        } else {
            xlim <- range(values)
        }

        boxplot(values, horizontal = TRUE, yaxt = "n", col = col, xlab = xlab, main = main, ylim = xlim, ...)
        if(zero) {
            abline(v = 0, col = "black", lty = 3, lwd = 1)
        }
    }

@@

    ## Which will be the centre plot?
    centre <- ceiling(n/2)
    ## Plot everything
    op <- par(mfrow = c(1, n+1))
    ## Plot the first panel
    par(bty = "n", mar = c(5, 14, 4, 0) + 0.1)
    empty <- matrix(0, ncol = length(results[[1]]$disparity))
    colnames(empty) <- names(results[[1]]$disparity)[rank]
    boxplot(empty, horizontal = TRUE, las = 2, col = "white", ylim = c(0,0), xaxt = "n", border = "white")
    for(i in 1:n) {
        ## Is position?
        is_position <- names(results)[[i]] == "position"
        par(mar = c(5, 0, 4, 0) + 0.1, bty = "o")
        one.plot(results[[i]], col = col[rank], centre = is_position && centre.position, main = ifelse(i == centre, main, ""), xlab = names_list[i], zero = FALSE, absolute = absolute, rank = rank, xlim = xlim)
    }
    par(op)
}
wrap.plot(results$super_results_between,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"))
```

@@

<details>
  <summary>[Click to expand] Same but zooming on the (0,5) interval (to get less of the flamingo effect)
</summary>

```{r, echo = FALSE, fig.height = 8, fig.width = 8}
wrap.plot(results$super_results_between,
          xlim = c(0,6),
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[groups on phylogeny] superorders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"),
          centre.position = TRUE, absolute = TRUE)
```
</details>

```{r, fig.height = 16, fig.width = 8, echo = FALSE}
wrap.plot(results$order_results_between,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"))
```

<details>
  <summary>[Click to expand] Again, same but with some zoom (0,15)
</summary>

```{r, echo = FALSE, fig.height = 16, fig.width = 8}
wrap.plot(results$order_results_between,
          xlim = c(0, 15),
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[groups on phylogeny] orders VCV\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"),
          centre.position = TRUE, absolute = TRUE)
```
</details>

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE, eval = TRUE}
## Plot the results
wrap.plot(results$super_results_phylo,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on phylogeny] superorders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"))
```

@@

## Plot the results
wrap.plot(results$order_results_phylo,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on phylogeny] orders species\nonto the phylogenetic VCV",
          names = c("elaboration", "innovation"))
```

@@

```{r, fig.height = 8, fig.width = 8, echo = FALSE}
## Plot the results
wrap.plot(results$super_results_within,
          col = rainbow(n.subsets(superorder_shapespace)-1),
          main = "[species on group] superorders species\nonto their group's VCV",
          names = c("elaboration", "innovation"))
```

@@

```{r "before last", fig.height = 16, fig.width = 8, echo = FALSE}
## Plot the results
wrap.plot(results$order_results_within,
          col = rainbow(n.subsets(order_shapespace)-1),
          main = "[species on group] oreders species\nonto their group's VCV",
          names = c("elaboration", "innovation"))
```
